---
title: "How to handle NA in PoPLR?"
date: Tue 10 Jan 2023 09:56:17 AWST
output:
  html_document:
    toc: true
    theme: cerulean
    number_sections: true
---

```{r}
# Mon 26 Dec 2022 11:57:08 AEDT
source('poplr.r')
```

# Post ARVO submission

```{r make-series-from-res}
#
# Convert data into a good format for PoPLR.
#
# @param res List indexed as [[i_eye]][[i_rep]][[i_visit]][, c(..., "npres", "final")]
# @param visits Take the first `visits` measurements for each eye-rep.
# @param field Either "final" or "npres"
#
# @return List of final thresholds as [[i_eye]][[i_rep]][location, visit].
#         Pad any locations that do not have all visits with NA at the early visits.
#         Assumes that once a location has commenced testing, it remains tested.
#
make_series <- function(res, visits = 1:5, field = "final") {
    lapply(res, function(eye) {
        lapply(eye, function(rep) {
            n <- nrow(rep[[tail(visits, 1)]])    # max number of locations for this rep
            d <- sapply(visits, function(visit) {
                t <- rep[[visit]][, field]
                if (length(t) < n)
                    t <- c(t, rep(NA, n - length(t)))
                t
            })
        })
    })
}
```


```{r}
# Tue 10 Jan 2023 09:04:09 AWST
source('poplr.r')
load('arrest6.stable.2.Rdata')
a <- make_series(res[[3]], 1:5)
b <- a[[25]][[1]]

b

z <- b[,5] > 16
zz <- b[,5] > 16 & 1:nrow(b) <= 52

PoPLR(b[z, ], w = FALSE)
PoPLR(b[zz, ], w = FALSE)

PoPLR(b[53:56, ], w = FALSE)
PoPLR(b[53:56, 2:5], w = FALSE)

PoPLR(matrix(c(
    30, 30, 30, 30, 30,
    NA, NA, 20, 15, 10), nrow = 2, byrow = TRUE), w = FALSE)

PoPLR(matrix(c(
    30, 30, 30, 30, 30,
    NA, 25, 20, 15, 10), nrow = 2, byrow = TRUE), w = FALSE)

PoPLR(matrix(c(
    30, 30, 30+10, 30, 30,
    NA, NA, 20+10, 15, 10), nrow = 2, byrow = TRUE), w = FALSE)

PoPLR(matrix(c(
    30, 30, 30+10, 30, 30,
    NA, 25, 20+10, 15, 10), nrow = 2, byrow = TRUE), w = FALSE)
```

Using PoPLR that does not correct for repeated perms in sequences with NAs:
 
|  Run                             | squish NAs   | diff perms    | selected perms |
|:---------------------------------|-------------:|--------------:|----------------:|
| PoPLR(b[z, ], w = FALSE)         |  0.01680672  |   0.06666667  | 0.06722689  |
| PoPLR(b[zz, ], w = FALSE)        |  0.3319328   |   0.3375      | 0.3319328   |
| PoPLR(b[53:56, ], w = FALSE)     |  0.1134454   |   0.1041667   | 0.06521739  |
| PoPLR(b[53:56, 2:5], w = FALSE)  |  0.06521739  |   0.1041667   |  0.06521739 |
| PoPLR( matrix(c(                 |              |               |             |
|  &#12;  30, 30, 30, 30, 30,          |               |               |           |
|  &#12;  30, NA, 20, 15, NA), ...)    | 0.01260504    | 0.008333333   |           |
|PoPLR(matrix(c(                       |               |               |           |                        
|  &#12;  30, 30, 30, 30, 30,          |               |               |           |        
|  &#12;  30, NA, 20, NA, 10), ...)    | 0.02941176    | 0.008333333   |           |
|                                      |               |               |           |
|PoPLR(matrix(c(                       |               |               |           |
| &#12; 30, 30, 30, 30, 30,            |               |               |           |
| &#12; NA, NA, 20, 15, 10), ...))     |               |               | 0.0210084 |
|PoPLR(matrix(c(                       |               |               |           |
| &#12; 30, 30, 30, 30, 30,            |               |               |           |
| &#12; NA, 25, 20, 15, 10), ...))     |               |               | 0.004202  |
| PoPLR( matrix(c(                     |               |               |           |
|  &#12;   30, 30, 30+10, 30, 30,      |               |               |           |
|  &#12;   NA, NA, 20+10, 15, 10), ...)| 0.01260504    | 0.008333333   | 0.0210084 |
|PoPLR(matrix(c(                       |               |               |           |
|  &#12;  30, 30, 30+10, 30, 30,       |               |               |           | 
|  &#12;  NA, 25, 20+10, 15, 10), ...) |               |               | 0.02941176|


The principle behind each method:

 * squish - just ignore the NAs and leave them in for everything (in effect altering/stretching the 
            gaps between xs for some perms)
 * diff perms - use different perms for shorter sequences so NAs are left out (worried about GVE here)
 * selected perms - use one set of perms, but for rows with NAs, leave out perms that do not have NAs 
                    just at either end (ie we don't know what happened at that visit, so cannot 
                    include it in a perm)