---
title: "Incorporating Events Into PoPLR"
output:
  html_document:
    toc: true
    theme: cerulean
    number_sections: true
---
# Scenario

## Data
We have a series of visual fields for an eye measured at visits $1..n$ 
with ARREST. Let $m_{i} \in\{R, Y, 17..40\}$ be the measured visual threshold at visit $i \in [1,n]$.
We use $R$ for 'Red' (perimetrically blind), $Y$ for 'Yellow' which is
less than 17 dB but greater than $R$, and 
$17..40$ as dB thresholds (also known as $G$ or 'Green').

## Task
There are two main progression calling scenarios.
\begin{enumerate}
    \item Clinical Trial, where as soon as someone is called progressing, they 
    leave the trial. Typically $n$ is small like 8 or 10.

    \item Clinical Practice, where when progression is identified, it is treated and monitored,
    hoping that progression will stop or the rate will slow. In this scenario, $n$ can 
    span many visits, with different interventions at different time.
\end{enumerate}

Perhaps common to these two scenarios is the question 
\begin{quote}
    is this patient progressing at the same rate as last visit?
\end{quote}
(Assuming that stable = progressing at a rate of zero.)


## PoPLR only good for diagnosis

Because PoPLR has H0 as ``rate is 0'', I can't see how it can be used to answer the more 
general question with H0 as ``rate is $\beta$'' and H1: ``rate $\ne beta$''. 

As per ARVO 2023 submission, PoPLR works well in the clinical trial scenario as it is 
good for diagnosis, but does not give a rate, therefore is not that useful for monitoring.

It would have to be paired with a method for finding a rate (PLR, SPROG, ...).


## PoPLR and ARREST

Using PoPLR for diagnosis with ARREST works well 
for locations that are green for all visits $1..n$.
For locations that are added and are green at Visit $n$, it also seems to work, but I need 
to do some work on the repeated counting of sub-sequences with NAs in the permutations.

What to do with the Green to Yellow (G2Y) and Yellow to Red (Y2R) events?

Ideally we want them to contribute to the PoPLR ``probability that this is stable'' p value.
In the hypothesis testing framework like PoPLR we would have
\begin{itemize}
    \item H0: Event G2Y/Y2R is due to measurement noise and does not indicate progression.
    \item H1: Event G2Y/Y2R is due to progression.
\end{itemize}

# Idea 1

For a subsequence $\{m_a, \ldots, m_b = Y\}$, where all $m_i$ are Green 
for $1 \le a \le i < b \le n$,
what is the chance that $m_b \ge 17$?

If we do PLR on $\{m_a, \ldots, m_{b-1}\}$ and use that to predict $m_b$, say $m_b^\prime$, 
how far off Yellow is that?

But if slope of PLR is shallow, wont get near Yellow, and if it is steep then PoPLR should 
have already picked it up (ie already had low p-value in visit $< b$).

Thus we should:
\begin{itemize}
    \item (at least) include the past green section ($m_a\ldots m_{b-1}$) in PoPLR; and 
    \item if the field ``is stable'' (high p value) look at the difference between the mean of the 
          Green and 16 dB.
\end{itemize}

Perhaps the question is: what is the probablility of getting Yellow if your true threshold
is stable at ``mean of Green'' $M = \mbox{mean}(\{m_a,ldots,m_{b-1}\})$ (or whatever true 
threshold would be most likely to give you $M$).

To get a G2Y you need to get a ZEST threshold  $\le 16$, not see 16 twice.

## Model 1

```{r Henson}
# Mon  9 Jan 2023 17:43:07 AWST
pr_seeing <- function(x, tt, fpr=0, fnr=0) {
    sd <- min(6, exp(3.27 -0.081 * tt))
    return (fpr + (1 - fpr - fnr)*(1 - pnorm(x, tt, sd)))
}
```

```{r model1}
tts <- 17:30
cbind(tts, 
    round(sapply(tts, function(tt) (1 - pr_seeing(16, tt))^2), 3),
    round(sapply(tts, function(tt) (1 - pr_seeing(16, tt, fpr = 0.15, fnr = 0.03))^2), 3)
)
```


# Aside: hetroscedasticity and weighted regression

Shouldn't we be using weighted regression?
Actually, maybe in the Green range it is all OK - hetro. enough.

Allison points out that it is OK in any sequence when $abs(m_n - m_1)$ is smallish: that is, 
the variance of the errors are roughly equal (all bad, all good, etc).

\end{document}